---
title: "Week_1_Basics"
author: "Laura Palacios"
date: '2022-09-28'
output: html_document
---
Hello! Welcome to Laura and Sam's R101 teaching sessions! Today we're going to go over some of the R basics.  
```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

Installing packages
```{r}

```

Installing libraries
```{r}

```

Coding basics

i.	R variables
1.	logical
2.	numeric
3.	integer
4.	complex
5.	character
6.	raw
```{r}

```


ii.	How to assign a variable
```{r}

```

iii.	Checking variable type
```{r}

```

iv.	Base operators

1.	Arithmetic
2.	Math – min/max/sum
3.	Equal or unequal
4.	And vs Or 
5.	%in%
```{r}

```

v.	Base Functions
```{r}

```

1.	Print
```{r}

```

2.	Concatenate
```{r}

```

3.	Is.na()
```{r}

```

vi.	Data Classes: String vs list vs data frame 
```{r}

```

vii.	Indexing




ADJECTIVE
ANIMAL
PLACE
VERB_Present
NOUN
VERB_Past
ADVERB
EXCLAMATION
FRIEND_NAME
CELEBRITY
SILLY_WORD
NUMBER
ADJECTIVE2
NOUN_Plural
ADJECTIVE3
VERB_Past2
BODY_PART_Plural
VERB_
NOUN_PLURAL_2
VERB_ING
NOUN_PLURAL_2
ADJECTIVE4








```{r}
print('The Best Halloween Party Ever
Last night I went to the most ' _________ Halloween
party! The invitation was delivered by ____________
and told me to go to the deep dark ______________
and _________ all the way to the very top of the
spooky __________. I __________ the doorbell
__________. ___________! My friend __________ answered
the door dressed up as _____________ and said
“______________!” There were _________ different 
costumes, including _____________ ghouls and mummified
_____________. The music was loud and ___________, so
I ______________ until my __________ hurt. There were
also delicious treats to __________, like candy corn and
chocolate-covered ______________, but my favourite part of
the party was the pumpkin _____________ and bobbing for
______________. I can’t wait until next Halloween—
it will be even more ______________!')