---
title: "Week_1_Basics"
author: "Laura Palacios and Samantha Olivares-Mejia"
learner:
date: '2022-09-28'
output: html_document
---
```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

# R101 - Week 1

Hello! Welcome to Laura and Sam's R101 teaching sessions! Today we're going to go over some of the R basics.

## Navigating R studio

## Projects vs Single Scripts

## RMD format
```{r}
# Sam, 
#Can you briefly mention that symbols (# and <>) indicate output formatting. I think we can have a whole separate section on this, but I don't want people to be confused. OR, if you hate the idea, we can remove the formatting?
```


## Installing packages

## Installing libraries

## R Variables

### What is an R variable?
A variable is a way to store a piece of information. It can later be used to reference or manipulate.


### How does it work?
The information is first named, then assigned a value. The value assignment can either be done directly by the user or programmatically assigned.

**TO DO** Below, write the most commonly used way of assigning a value in R.

How to assign a value:

*Hint: Python uses an =.*

*Hint: Example seen below.*


### Variable names
There are some rules to follow when assigning names. These include:
  1. Easy to understand
  2. No spaces or special characters.
  3. Starts with a letter. 
  4. Letter is is followed by any letter, number, underscore, or period.
  5. No reserved words (reserved words are used in arguements and may cause computer confusion if used as a variable name)  Here is a list: https://rdrr.io/r/base/Reserved.html
  
**TO DO** Write a possible name for a variable which contains bird data

Bird data variable name:

*Hint: One example could be frequent_flyers*


### Variable practice
To start, we'll assign some simple variables for the different types of data that R uses.
<br/><br/>


#### 1. Numeric data
Variables can be assigned any number.

Here is an example:
```{r}
# R can hold whole numbers
Whole_Number <- 1
# R can hold multiple decimal places
Decimal <- 7.5895892
```
Run this code chunk (click the green arrow in the top right of the chunk).

Notice the environment (top right) now contains your variables and their assigned values.

  
**TO DO** Assign the variables in the code block below
```{r}
# Assign a negative number 
Negative_Number <- 9999
  
# Assign any other number
Any_Number <- 9999
```
*Tip: Notice that when a variable is properly assigned it changes color*


**TO DO** Assign a new number to the Decimal variable
```{r}
# Assign a new number
Decimal <-9999
```
Notice that the value has changed in your environment. 


**TO DO** What value would appear if we ran the first code (Decimal <- 7.5895892) again?

Answer:

*Hint: The value is always overwritten by the last code block, regardless of order in script.*
<br/><br/>


#### 2. Integers
Integers are a numeric value with no decimal points.

A suffix of L indicates that the value is an integer.Example below.
```{r}
# Assign an integer - example
Integer <- 1L
```


**TO Do** Assign your own integer
```{r}
# Assign an integer
My_Integer <- 9999
```
*Tip: Integers have a numeric color when properly assigned*
<br/><br/>


#### 3. Character data
Character data  can include words, sentences, filepaths, etc. It is stored and treated as a block of text.

Example:
```{r}
# Notice that single quotes are used to indicate the string component
Magical_Phrase <- 'Hello World'
```
Notice that the character data also occurs in your environment.

**TO DO** Lets try assigning some:
```{r}
# Assign a descriptive word
Adjective <- '9999'

# Assign your favorite animal
Animal <- '9999'

# Assign a celebrity name
Celebrity <- '9999'

# Assign anything!
Thing <- '9999'

# Assign a different thing
# Notice the 2 indicates that this is a different variable and the original Thing is not overwritten.
Thing2 <- '9999'
```

**TO DO** Can I use double quotes?
```{r}
#Assign a happy phrase in double quotation marks (")
Exclamation <- "9999"
```
*Hint: Yes! Single or double quotes may be used.*

**TO DO** Can I use more than one word?
```{r}
# Write the favorite activity you did last week
# Example: Activity <- "swam in the pool" 
#or
#Example: Activity <- "Ran on a trail"
Activity <- "9999"
```
*Hint: The quotation marks should be placed around the entire phrase*
<br/><br/>


#### 4. Complex
Complex data handles imaginary numbers in R

Imaginary numbers are indicated with an i.

**To Do** Replace this imaginary animal with an imaginary number such as 2i.
```{r}
# Attribute Imaginary with a complex number. 
Imaginary <- 'Unicorn'
```
*Hint: Complex values do not require quotation marks, proper formatting will change the color to the numeric color*
<br/><br/>


#### 5. Logical/ Boolean
Logical data is True or False.

Example:
```{r}
# Assign as a TRUE logical value.
Logic_True <- TRUE
```
Notice that the TRUE value is all caps and the color is the same as the complex and numeric variables.

**To Do** Create a false variable
```{r}
# Assign a false logical value
Logic_False <- 9999
```
*Tip: Autofill can be used. After typing "FAL" a dropdown will appear. Press down arrow and enter OR click to select*
<br/><br/>


#### 6. Raw
Converts other data types to raw bytes. This is not a common type.

Example:
chartoRaw() and intToBits() are used to convert a value to a raw variable type.
```{r}
# Character to raw example
Raw_Character <- charToRaw("Hello World")
# Integer to bits example
Raw_Int <- intToBits(7L)
```


## R Data Structures

### What are data structures?
Data structures are storage components for holding multiple values.
<br/><br/>


#### 1. Vectors 
A vector is a collection of data values, such as characters, logical, integer, or numeric. It can be treated like a single column on a table.

Example. Here is a character vector. Use the notation c() (column) to create the collection of values.
```{r}
# Create a character vector
Character_Vector <- c('Snoopy', 'Charlie Brown', 'Linus')
```
*Tip: Notice the environment show characters [1:3] (count of data) then lists the values*


**To do** Create an Atomic Vector, a vector that consists of one value, containing your favorite location. 
```{r}
# Assign a place
Place <- '9999'
```
*Hint: We made atomic vectors in the Variables section*


**To do** Create a numeric vector
```{r}
# Create a numeric vector
Number_Vector <- 9999
```
*Hint: separate values with commas and use the c() notation*


Vectors can use a combination of data types.
**To do**
Create a vector with 3 components of different data types.
```{r}
# Create a mixed vector
Mixed_Vector <- 9999
```
*Hint: Try a mix character, integer, numeric, or logical values*


Data structures, including vectors, can also reference previously created variables by using their names.
```{r}
# Create a mixed vector of different data types
My_Vector <- c(9999, 9999, 9999r)
```
*Tip: Notice that this becomes a chr (character) list (shown in the environment). R will automatically determine an appropriate format unless otherwise specified by you*
<br/><br/>


#### 2. Lists
A list is a collection of data values, such as characters, logical, integer, or numeric. 
So why is this different than a vector?

Lists are ordered, recursive (loopable) and multidimensional!
(They are more complex so you can do more things!)


Here is an example of a simple list
```{r}
# Create a short list
List_Example<- list('test', 1, 2, 3)
```
Notice that the list appears in the data section of the environment.


Is it possible to create a list of other items?

**TO DO:** Create a list of your vector (My_Vector), List Example(List_Example), and one other atomic vector you previously created.
```{r}
# Create a list of lists
List_Of_Lists <- 9999
```
*Hint: Use the function list() to create a list*
<br/><br/>


#### 3. DataFrames
A data frame is a two dimensional way to store data, similar to a sheet in Microsoft Excel or Google sheets. We can also think of it as two or more vectors.

Example: Here is a simple dataframe created in R.
```{r}
# Create a vector for the first column
Song <- c("Rings", "Birds", "Hens", "Doves", "Partridge in a Pear Tree")
# Create a vector for the second column (same length)
Count <- c(5,4,3,2,1)

#Create the dataframe
Example_Dataframe <- data.frame(Song,Count)
```
*Tip: click on the data frame from the environment to open a new tab and view the sheet*
*Tip: press the down arrow in the environment to see a summary of data frame contents*


Please note. For most of SWNC work, we will be importing data frames (such as CSV files) rather than creating our own. This will be covered in another session.
<br/><br/>


#### 4. Matrices
A matrix is a set of rows and columns which organizes values of the same data type.

Mathmatical functions can be applied to matrices for complex analyses.
Here is an example matrix created in R:
```{r}
# The matrix creation requires the values (1-9), number of rows and/or number of columns)
Matrix_example <- matrix(c(1:9), nrow = 3, ncol = 3)
```
*Tip: Click on the matrix example in the environment to open a new tab and view all data*
<br/><br/>


#### 5. Arrays
Arrays are objects with two or more dimensions. Again, these are used for more complicated analyses in R. 

Example, a three dimensional array can be thought of as three stacked matrices, which will have a number of columns, rows and height.
```{r}
# Create an example array
Array_Example <- array(0, dim = c(2, 2, 3))
# Show the array named "Array_Example" below the code block
print(Array_Example)
```
*Tip: The array is considered a "value" in the environment of Rstudio. You can easily view the array (or any other object) by using the print() function to show it below a code block. We will go into more detail of the print() function later.*
<br/><br/>


#### 6. Factors
Factors are used to store categorical (grouped) data  in a way that can be used for calculations or other machine learning processes.

Factors can be ordinal (they have a rank or values are ordered) or non-ordinal.
Here is an example:
```{r}
# Create the categorical vector (any order is okay)
day_vector <- c('afternoon','night','morning')

# Convert `day_vector` to a factor with ordered level
# Order = TRUE indicates that it is ordinal
# levels indicates the order. It is only required if the order = TRUE
Factor_example <- factor(day_vector, order = TRUE, levels =c('morning', 'afternoon','night'))

#Print the Factor example below.
print(Factor_example)
```
**To Do:** Fill in the blanks to create a non-ordinal factor with three values.
```{r}
# Create the vector
My_Vector <- c(9999)

#My_Factor <- factor(My_Vector, order = )
```



Checking variable type

Base operators
1.	Arithmetic
2.	Math – min/max/sum
3.	Equal or unequal
4.	And vs Or 
5.	%in%

Sam please find a way to attribute numbers A-E for the activity.
```{r}
Number_A <- 9999
Number_B <- 9999
Number_C <- 9999
Number_D <- 9999
Number_E <- 9999
```

Base Functions

Print

Concatenate

Is.na()

Indexing
Sam - please find a way to use activity[1]. and assign to "Verb" for the activity.
```{r}
Verb <- Activity[1]
```







```{r, include = FALSE}


######################### End of Document. Stop here#########################################


```

























```{r, include = FALSE}


##################### I mean it, stop scrolling, you'll ruin the surprise. #######################


```







```{r, echo= FALSE}
print('The Best Halloween Party Ever') 
sprintf("Last night I went to the most %s Halloween party! The invitation was delivered by %s and it told me to go to the deep dark %s. I %s the doorbell %f times.My friend answered the door dressed up as %s and said '%s!' There were %f different costumes, including %f ghouls and %f mummified %ss. There were also delicious treats to eat like candy corn and %f candy bars, but my favourite part of the party was the pumpkin carving and bobbing for %ss. I can't wait until next Halloween!'", Adjective, Animal, Place, Verb, Number_A, Celebrity, Exclamation, Number_B, Number_C, Number_D,Thing, Number_E, Thing2)
```
 